Two-Dimensional Vesicle Hydrodynamics from Hydrophobic Attraction Potential

### 2021-12-29 doc/main_jfm.tex
working on revisions

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mobility Problem Boundary Integral Equation}
Following previous work of~\cite{Lukas19}, we use the velocity
representation of~\cite{pow-mir1987}. In particular, we write the
velocity as the sum of a double-layer potential and $N_b$-many Stokeslets and
rotlets
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega.
\end{align}
The double-layer potential is
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nnu}{|\rr|^2} \frac{\rr \otimes \rr}{|\rr|^2}
    \cdot \eeta(\yy) \, \dif s_\yy,
\end{align}
where $\rr = \xx - \yy$ and $\rho = |\rr|$. The Stokeslet and rotlets
centred at $\aa_i$ are
\begin{align}
  \SS(\xx,\aa_i) \cdot \FF_i &= \frac{1}{4\pi} \left(-\log |\rr| + 
    \frac{\rr \otimes \rr}{|\rr|^2}\right) \cdot \FF_i, \\
  \RR(\xx,\aa_i) T_i &= \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2} T_i,
\end{align}
respectively, where $\rr = \xx - \aa_i$. The Stokeslet is torque-free
and has force $\FF_i$ while the rotlet is force-free and has torque
$T_i$. Therefore, the velocity~\eqref{eq:velocity} satisfies the total
force~\eqref{eq:force} and torque~\eqref{eq:torque} conditions if the
double-layer potential $\DD[\eeta]$ is force- and torque-free.  Matching
the limit of~\eqref{eq:velocity} with the rigid body motion, and
imposing that $\DD[\eeta]$ is force- and torque-free, the density
function $\eeta$, translational velocity $\vv_i$, and angular velocity
$\omega_i$ satisfy
\begin{alignat}{3}
  \nonumber
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) \\
  \label{eq:SKIE}
    + \sum_{j=1}^{N_b} &
    \left(\SS(\xx,\aa_j) \cdot \FF_j + \RR(\xx,\aa_j) T_j\right),
    \quad &&\xx \in \Gamma_i,\: i=1,\ldots,N_b, \\
  \label{eq:mobility1}
  \int_{\bd\Gamma_i} \eeta \cdot \nnu_i \, \dif s &= {\bf 0}, 
  &&i = 1,\ldots,N_b, \\
  \label{eq:mobility2}
  \int_{\bd\Gamma_i} \eeta\times(\xx-\aa_i)^\perp \cdot \nnu_i \, \dif s &= 0,
  &&i = 1,\ldots,N_b.
\end{alignat}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Main Theoretical Result: Calculating the Hydrophobic Force}
\label{subsec:calculating_force}
Once \eqref{eq:screenedSKIE} has been solved for $\sigma$, we need to
evaluate the integrals~\eqref{eq:hydrophobicAttraction} which are the
HAP forces and torques. These integrals involve the
stress~\eqref{eq:stress} which contains a singular integral for the
gradient of the double-layer potential. To avoid singular integrals, we
first define
\begin{align}
  \label{eq:uidef}
  v_i(\xx) = u(\xx) - u_i(\xx),
\end{align}
where
\begin{align}
  u_i(\xx) &= \frac{1}{2\pi} \int_{\Gamma_i} \left(\pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right)\right) 
    \sigma(\yy) \, \dif s_\yy, \quad \xx \in \mathbb{R}^2.
\end{align}
That is, $v_i(\xx)$ is the double-layer potential~\eqref{eq:HAP_DLP}
with $\Gamma_i$ excluded from $\bd\Omega$. Having defined $v_i$, we prove
\begin{theorem}
\begin{align}
  \label{eq:recipforcetorque}
  \FF_i^{\mathrm{hydro}} = \int_{\Gamma_i} \JJ_i \,\dif s,\quad
  T_i^{\mathrm{hydro}}    = \int_{\Gamma_i} 
    (\xx - \aa_i)^{\perp} \cdot \JJ_i  \,\dif s,
\end{align}
where
\begin{align}
  \label{eq:jumpstress1}
  \JJ_{i} = 2\gamma\rho^{-1} \sigma v_i \nnu + 
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{ds} \nnu -
    2\gamma\rho \frac{d\sigma}{ds} \frac{dv_i}{d\nnu} \ttau.
\end{align}
\end{theorem}
The symbols $\ttau$ and $\frac{d}{ds}$ are the unit tangent and arc
length derivative for $\Gamma_i$, respectively (Figure
\ref{fig:figure0}a.) The result is valid for any smooth particle shape
or boundary condition. The advantage of
using~\eqref{eq:recipforcetorque} over
using~\eqref{eq:hydrophobicAttraction} is that the components $\sigma$,
$v_i$, $\frac{d\sigma}{ds}$, and $\frac{dv_i}{ds}$ of $\JJ_i$ are smooth
functions, whereas the components of \eqref{eq:stress} are singular
integrals.

@@

To prove \eqref{eq:recipforcetorque}, let 
\begin{align*}
  {\bf T} &=
    \SS(u_i,u_i) +(\SS(u_i,v_i) +\SS(v_i,u_i)) +\SS(v_i,v_i) \\
  &= {\bf T}_1 + {\bf T}_2 + {\bf T}_3
\end{align*}
where we introduce the bilinear form
\begin{equation}
\label{eq:Tsplit}
\SS(u,v)
=  \gamma\rho^{-1} uv {\bf I} + \gamma\rho \nabla u \cdot \nabla v {\bf I} - 2 \gamma \rho \nabla u \nabla v^T.
\end{equation}
Using the fact that $u$,  $u_i$, and $v_i$ solve the screened Laplace
equation~\eqref{eq:SL}, and that $\mathbf{T}$, $\mathbf{T}_j$, $j = 1,
2, 3$ are symmetric, it is straightforward to verify that 
  \begin{equation}
    \label{eq:decompdivfree}
    \nabla \cdot {\bf T}_j = 0, \quad
    \nabla \cdot ((\xx-\aa_i)^{\perp} \cdot {\bf T}_j) = 0, \quad j = 1, 2, 3.
  \end{equation}

@@

Let $U_i$ be the interior of the particle indexed by $i$. For $\xx_0
\in \Gamma_i$ and an arbitrary function $g(\xx)$, the notation
\begin{align}
  \jump{g}(\xx_0) = \lim_{\substack{\xx \to \xx_0 \\ \xx \in U_i^c}}g(\xx)  - 
                    \lim_{\substack{\xx \to \xx_0 \\ \xx \in U_i}}g(\xx),
\end{align}
denotes the jump of the limits of $g(\xx)$ taken from the outside to
the inside of $\Gamma_i$.
%\begin{align}
%\lim_{\xx \to \xx_0^+}g(\xx),\quad 
%\lim_{\xx \to \xx_0^-}g(\xx), \quad
%  \jump{g}(\xx_0) = \lim_{\xx \to \xx_0^+}g(\xx)  - \lim_{\xx \to \xx_0^-}g(\xx)
%\end{align}
%denotes the limit taken from points in $U_i^c$,  the limit taken from
%points in $U_i$, and the jump, respectively.  
\begin{lemma}
\begin{align}
  \label{eq:prejump}
  \FF_i^{\mathrm{hydro}} = \int_{\Gamma_i} \jump{{\bf T}_2 \cdot \nnu}  \, \dif s,\quad
  T_i^{\mathrm{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot
  \jump{{\bf T}_2 \cdot \nnu} \, \dif s.
\end{align}
\end{lemma}
\begin{proof}
To show~\eqref{eq:prejump}, we expand \eqref{eq:hydrophobicAttraction} as
\begin{align}
  \FF_i = \int_{\Gamma_i} \left({\bf T}_1 \cdot \nnu + 
    {\bf T}_2 \cdot \nnu + {\bf T}_3 \cdot \nnu \right) \,\dif s.
\end{align}
By~\eqref{eq:uidef} and~\eqref{eq:decompdivfree}, we have that $u_i$ is
smooth and $\nabla \cdot {\bf T}_1 = \mathbf{0}$ in $\mathbb{R}^n
\setminus U_i$. Similarly, $v_i = u - u_i$ is smooth and $\nabla \cdot
{\bf T}_3 = \mathbf{0}$ in $U_i$. By the divergence theorem,  
\begin{align}
  \int_{\Gamma_i}  {\bf T}_1 \cdot \nnu \,\dif s
  = -\int_{\mathbb{R}^n \setminus U_i} \nabla \cdot {\bf T}_1 \,\dif \xx = \mathbf{0},\quad
    \int_{\Gamma_i}  {\bf T}_3 \cdot \nnu\,\dif s
  = \int_{U_i} \nabla \cdot {\bf T}_3 \,\dif \xx = \mathbf{0}.
\end{align}
Finally, $u_i$ and $v_i$ are smooth and $\nabla \cdot {\bf T}_2= 0$ in $U_i$. This gives
\begin{align}
  \mathbf{0} = \int_{U_i} \nabla \cdot {\bf T}_2 \, \dif \xx = -\int_{\Gamma_i}  
    ({\bf T}_2 \cdot \nnu)^-\,\dif s,
\end{align}
where the superscript denotes the limit taken from in $U_i$.  
Combining the above gives the first equation in \eqref{eq:prejump}.
The argument for the second equation in \eqref{eq:prejump} is identical. 
\end{proof}

We have the following jump relations for \eqref{eq:uidef}: 
\begin{equation}
\label{eq:DLjump}
\jump{u_i} = \sigma, \quad
\jump{v_i} = 0, \quad
\jump{\nabla u_i} = \frac{\dif \sigma}{\dif s}\ttau,\quad
\jump{\nabla v_i} = \mathbf{0},
\end{equation}
on $\Gamma_i$ (see, e.g. \cite{KlBaGrON13}). Therefore,
\begin{align*}
  \jump{{\bf T}_2 \cdot \nnu}   &= \jump{\SS(u_i,v_i) \cdot \nnu +
    \SS(v_i,u_i) \cdot \nnu} \\
  &= \Big[\!\Big[ \left( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla
  u_i^T\right) \cdot \nnu \Big]\!\Big] \\
%  &= \jump{\left( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
%- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla
%  u_i^T\right) \cdot \nnu}\\
&= 2\gamma\rho^{-1} \sigma v_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dv_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dv_i }{d\nu} \ttau = \JJ_i.
\end{align*}
Combining this with~\eqref{eq:prejump} gives~\eqref{eq:recipforcetorque}
as required.

@@

To calculate $U$, we let $u_+(\mathbf{x})$ and $u_-(\mathbf{x})$ 
be the tangential velocity of the outer, respectively inner, leaflet.  Then 
\begin{equation}
U = \frac{1}{L}\int_{C_M}u_+(\mathbf{x})(1 - \delta \kappa) - v_-(\mathbf{x})(1 +  \delta \kappa) \, ds,
\end{equation}
where $ \delta $ is the distance from the leaflet centres to the midplane and $\kappa$ is the curvature. 
The factors $\pm  \delta \kappa$ are needed to project the leaflet velocities, defined on the particle centres,
onto the midplane curve. 
Finally, the function $s(t) = \int_0^t dU $ records the distance one leaflet has
slid past the other. Figure~\ref{figure5}b plots the sliding distance for various shear
rates.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mobility Problem Boundary Integral Equation}
Following previous work of~\cite{Lukas19}, we use the velocity
representation of~\cite{pow-mir1987}. In particular, we write the
velocity as the sum of a double-layer potential and $N_b$-many Stokeslets and
rotlets
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega.
\end{align}
The double-layer potential is
\begin{align}
  \DD[\eeta](\xx) = \frac{1}{\pi} \int_{\bd\Omega} 
    \frac{\rr \cdot \nnu}{|\rr|^2} \frac{\rr \otimes \rr}{|\rr|^2}
    \cdot \eeta(\yy) \, \dif s_\yy,
\end{align}
where $\rr = \xx - \yy$. The Stokeslet and rotlets centred at $\aa_i$
are
\begin{align}
  \SS(\xx,\aa_i) \cdot \FF_i &= \frac{1}{4\pi} \left(-\log |\rr| + 
    \frac{\rr \otimes \rr}{|\rr|^2}\right) \cdot \FF_i, \\
  \RR(\xx,\aa_i) T_i &= \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2} T_i,
\end{align}
respectively, where $\rr = \xx - \aa_i$. The Stokeslet is torque-free
and has force $\FF_i$ while the rotlet is force-free and has torque
$T_i$. Therefore, the velocity~\eqref{eq:velocity} satisfies the total
force~\eqref{eq:force} and torque~\eqref{eq:torque} conditions if the
double-layer potential $\DD[\eeta]$ is force- and torque-free.  Matching
the limit of~\eqref{eq:velocity} with the rigid body motion, and
imposing that $\DD[\eeta]$ is force- and torque-free, the density
function $\eeta$, translational velocity $\vv_i$, and angular velocity
$\omega_i$ satisfy
\begin{alignat}{3}
  \nonumber
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) \\
  \label{eq:SKIE}
    + \sum_{j=1}^{N_b} &
    \left(\SS(\xx,\aa_j) \cdot \FF_j + \RR(\xx,\aa_j) T_j\right),
    \quad &&\xx \in \Gamma_i,\: i=1,\ldots,N_b, \\
  \label{eq:mobility1}
  \int_{\bd\Gamma_i} \eeta \cdot \nnu_i \, \dif s &= {\bf 0}, 
  &&i = 1,\ldots,N_b, \\
  \label{eq:mobility2}
  \int_{\bd\Gamma_i} \eeta\times(\xx-\aa_i)^\perp \cdot \nnu_i \, \dif s &= 0,
  &&i = 1,\ldots,N_b.
\end{alignat}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Main Theoretical Result: Calculating the Hydrophobic Force}
\label{subsec:calculating_force}
Once \eqref{eq:screenedSKIE} has been solved for $\sigma$, we need to
evaluate the integrals~\eqref{eq:hydrophobicAttraction} which are the
HAP forces and torques. These integrals involve the
stress~\eqref{eq:stress} which contains a singular integral for the
gradient of the double-layer potential. To avoid singular integrals, we
first define
\begin{align}
  \label{eq:uidef}
  w_i(\xx) = u(\xx) - u_i(\xx),
\end{align}
where
\begin{align}
  u_i(\xx) &= \frac{1}{2\pi} \int_{\Gamma_i} \left(\pderiv{}{\nnu_\yy}
    K_0 \left(\frac{|\xx - \yy|}{\rho}\right)\right) 
    \sigma(\yy) \, \dif s_\yy, \quad \xx \in \mathbb{R}^2.
\end{align}
That is, $w_i(\xx)$ is the double-layer potential~\eqref{eq:HAP_DLP}
with $\Gamma_i$ excluded from $\bd\Omega$. Having defined $w_i$, we prove
\begin{theorem}
\begin{align}
  \label{eq:recipforcetorque}
  \FF_i^{\mathrm{hydro}} = \int_{\Gamma_i} \JJ_i \,\dif s,\quad
  T_i^{\mathrm{hydro}}    = \int_{\Gamma_i} 
    (\xx - \aa_i)^{\perp} \cdot \JJ_i  \,\dif s,
\end{align}
where
\begin{align}
  \label{eq:jumpstress1}
  \JJ_{i} = 2\gamma\rho^{-1} \sigma w_i \nnu + 
    2\gamma\rho \frac{d\sigma}{ds} \frac{dw_i}{ds} \nnu -
    2\gamma\rho \frac{d\sigma}{ds} \frac{dw_i}{d\nnu} \ttau.
\end{align}
\end{theorem}
The symbols $\ttau$ and $\frac{d}{ds}$ are the unit tangent and arc
length derivative for $\Gamma_i$, respectively (Figure
\ref{fig:figure0}a.) The result is valid for any smooth particle shape
or boundary condition. The advantage of
using~\eqref{eq:recipforcetorque} over
using~\eqref{eq:hydrophobicAttraction} is that the components $\sigma$,
$w_i$, $\frac{d\sigma}{ds}$, and $\frac{dw_i}{ds}$ of $\JJ_i$ are smooth
functions, whereas the components of \eqref{eq:stress} are singular
integrals.

@@

To prove \eqref{eq:recipforcetorque}, let 
\begin{align*}
  {\bf T} &=
    \SS(u_i,u_i) +(\SS(u_i,w_i) +\SS(w_i,u_i)) +\SS(w_i,w_i) \\
  &= {\bf T}_1 + {\bf T}_2 + {\bf T}_3
\end{align*}
where we introduce the bilinear form
\begin{equation}
\label{eq:Tsplit}
\SS(u,v)
=  \gamma\rho^{-1} uv {\bf I} + \gamma\rho \nabla u \cdot \nabla v {\bf I} - 2 \gamma \rho \nabla u \nabla v^T.
\end{equation}
Using the fact that $u$,  $u_i$, and $w_i$ solve the screened Laplace
equation~\eqref{eq:SL}, and that $\mathbf{T}$, $\mathbf{T}_j$, $j = 1,
2, 3$ are symmetric, it is straightforward to verify that 
  \begin{equation}
    \label{eq:decompdivfree}
    \nabla \cdot {\bf T}_j = 0, \quad
    \nabla \cdot ((\xx-\aa_i)^{\perp} \cdot {\bf T}_j) = 0, \quad j = 1, 2, 3.
  \end{equation}

@@

Let $U_i$ be the interior of the particle indexed by $i$. For $\xx_0
\in \Gamma_i$ and an arbitrary function $g(\xx)$, the notation
\begin{align}
  \jump{g}(\xx_0) = \lim_{\substack{\xx \to \xx_0 \\ \xx \in U_i^c}}g(\xx)  - 
                    \lim_{\substack{\xx \to \xx_0 \\ \xx \in U_i}}g(\xx),
\end{align}
denotes the jump of the limits of $g(\xx)$ taken from the outside to
the inside of $\Gamma_i$.
%\begin{align}
%\lim_{\xx \to \xx_0^+}g(\xx),\quad 
%\lim_{\xx \to \xx_0^-}g(\xx), \quad
%  \jump{g}(\xx_0) = \lim_{\xx \to \xx_0^+}g(\xx)  - \lim_{\xx \to \xx_0^-}g(\xx)
%\end{align}
%denotes the limit taken from points in $U_i^c$,  the limit taken from
%points in $U_i$, and the jump, respectively.  
\begin{lemma}
\begin{align}
  \label{eq:prejump}
  \FF_i^{\mathrm{hydro}} = \int_{\Gamma_i} \jump{{\bf T}_2 \cdot \nnu}  \, \dif s,\quad
  T_i^{\mathrm{hydro}} = \int_{\Gamma_i} (\xx - \aa_i)^{\perp} \cdot
  \jump{{\bf T}_2 \cdot \nnu} \, \dif s.
\end{align}
\end{lemma}
\begin{proof}
To show~\eqref{eq:prejump}, we expand \eqref{eq:hydrophobicAttraction} as
\begin{align}
  \FF_i = \int_{\Gamma_i} \left({\bf T}_1 \cdot \nnu + 
    {\bf T}_2 \cdot \nnu + {\bf T}_3 \cdot \nnu \right) \,\dif s.
\end{align}
By~\eqref{eq:uidef} and~\eqref{eq:decompdivfree}, we have that $u_i$ is
smooth and $\nabla \cdot {\bf T}_1 = \mathbf{0}$ in $\mathbb{R}^n
\setminus U_i$. Similarly, $w_i = u - u_i$ is smooth and $\nabla \cdot
{\bf T}_3 = \mathbf{0}$ in $U_i$. By the divergence theorem,  
\begin{align}
  \int_{\Gamma_i}  {\bf T}_1 \cdot \nnu \,\dif s
  = -\int_{\mathbb{R}^n \setminus U_i} \nabla \cdot {\bf T}_1 \,\dif \xx = \mathbf{0},\quad
    \int_{\Gamma_i}  {\bf T}_3 \cdot \nnu\,\dif s
  = \int_{U_i} \nabla \cdot {\bf T}_3 \,\dif \xx = \mathbf{0}.
\end{align}
Finally, $u_i$ and $w_i$ are smooth and $\nabla \cdot {\bf T}_2= 0$ in $U_i$. This gives
\begin{align}
  \mathbf{0} = \int_{U_i} \nabla \cdot {\bf T}_2 \, \dif \xx = -\int_{\Gamma_i}  
    ({\bf T}_2 \cdot \nnu)^-\,\dif s,
\end{align}
where the superscript denotes the limit taken from in $U_i$.  
Combining the above gives the first equation in \eqref{eq:prejump}.
The argument for the second equation in \eqref{eq:prejump} is identical. 
\end{proof}

We have the following jump relations for \eqref{eq:uidef}: 
\begin{equation}
\label{eq:DLjump}
\jump{u_i} = \sigma, \quad
\jump{w_i} = 0, \quad
\jump{\nabla u_i} = \frac{\dif \sigma}{\dif s}\ttau,\quad
\jump{\nabla w_i} = \mathbf{0},
\end{equation}
on $\Gamma_i$ (see, e.g. \cite{KlBaGrON13}). Therefore,
\begin{align*}
  \jump{{\bf T}_2 \cdot \nnu}   &= \jump{\SS(u_i,w_i) \cdot \nnu +
    \SS(w_i,u_i) \cdot \nnu} \\
  &= \Big[\!\Big[ \left( 2\gamma\rho^{-1} u_i w_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla w_i \mathbf{I} 
- 2\gamma\rho \nabla u_i  \nabla w_i^T - 2\gamma\rho \nabla w_i \nabla
  u_i^T\right) \cdot \nnu \Big]\!\Big] \\
%  &= \jump{\left( 2\gamma\rho^{-1} u_i v_i \mathbf{I} + 2\gamma\rho \nabla u_i \cdot \nabla v_i \mathbf{I} 
%- 2\gamma\rho \nabla u_i  \nabla v_i^T - 2\gamma\rho \nabla v_i \nabla
%  u_i^T\right) \cdot \nnu}\\
&= 2\gamma\rho^{-1} \sigma w_i \nnu + 2\gamma\rho \frac{d\sigma }{ds}\frac{dw_i }{ds} \nnu
- 2\gamma\rho \ \frac{d\sigma }{ds} \frac{dw_i }{d\nnu} \ttau = \JJ_i.
\end{align*}
\todo[inline]{RR: Can you make sure the last line is correct? There was
an unbolded $\nu$ which I changed to a bolded $\nnu$.}
Combining this with~\eqref{eq:prejump} gives~\eqref{eq:recipforcetorque}
as required.

@@

To calculate $U$, we let $u_+(\mathbf{x})$ and $u_-(\mathbf{x})$ 
be the tangential velocity of the outer, respectively inner, leaflet.  Then 
\begin{equation}
  U = \frac{1}{L}\int_{C_M}\left(u_+(\mathbf{x})(1 - \delta \kappa) -
  u_-(\mathbf{x})(1 +  \delta \kappa)\right) \, ds,
\end{equation}
where $ \delta $ is the distance from the leaflet centres to the midplane and $\kappa$ is the curvature. 
The factors $\pm  \delta \kappa$ are needed to project the leaflet velocities, defined on the particle centres,
onto the midplane curve. 
Finally, the function $s(t) = \int_0^t dU $ records the distance one leaflet has
slid past the other. Figure~\ref{figure5}b plots the sliding distance for various shear
rates.
